Grade Retention and Mixed Race

### 2022-08-18 products/paper/main.qmd
Add a different start to intermarriage section

## The Role of Class

In addition to the aforementioned racial mechanisms that may drive the placement of biracial individuals, researchers must also pay careful attention to how socioeconomic background may affect the placement of biracial individuals within a racial hierarchy. Socioeconomic background plays a role in racial inequality more generally as we know that accounting for socioeconomic difference often reduces the magnitude of observed racial inequality, suggesting that some of the racial inequality we observe operates through class differences (CITATION). This issue is particularly important for biracial individuals because of potential class selection into interracial unions. The likelihood of being an interracial union varies by class background for at least some groups (CITATION), although it varies in complex ways and we know more about interracial marriages than other types of unions that may produce children. For example, black partners tend to be positively selected with regard to education into Black/White marriages while white partners are negatively selected (CITATION). According to status exchange theory (CITATIONS), this pattern may be common to all types of interracial marriages within an unequal status hierarchy, although the evidence of status exchange for other forms of marriages is weaker (CITATION). How this affects the relative resources of biracial individuals will depend then in some part on the relative strength of these phenomena for a given pairing, the gender balance in that pairing, and how the gender of parents affects outcomes.

>>>>

## Selection into Intermarriage

The placement of biracial individuals may also be affected by mechanisms that drive selection into interracial unions in the previous generation. Such selection may lead to differences on a variety of parental background characteristics between biracial individuals and their monoracial peers. How this selection will affect a biracial child's outcomes for a given type of pairing are complex because it depends on the strength and direction of that selection on each partner, its consistency across different types of unions, and how strongly the selected characteristic predicts outcomes in the next generation.

Most research on such selection has focused on socioeconomic characteristics. Status exchange theory predicts that given an unequal racial hierarchy, individuals from a lower status racial group will be positively selected on socioeconomic characteristics into intermarriage with a higher status group, while individuals from a higher status racial group will be similarly negatively selected into the same marriage (CITATION). Using education as the socioeconomic parameter, prior research has shown strong evidence of this pattern in the case of Black/White marriages in the US (CITATION). Evidence of this pattern for other forms of intermarriage is less clear (CITATION).

Alternatively,

---

The aforementioned mechanisms all relate to

In addition to the aforementioned racial mechanisms that may drive the placement of biracial individuals, researchers must also pay careful attention to how socioeconomic background may affect the placement of biracial individuals within a racial hierarchy. Accounting for socioeconomic difference often reduces the magnitude of observed racial inequality on a given outcome, suggesting that some of the racial inequality we observe operates through class differences (CITATION). This issue is particularly important for biracial individuals because of potential class selection into interracial unions. The likelihood of being an interracial union varies by class background for at least some groups (CITATION), although it varies in complex ways and we know more about interracial marriages than other types of unions that may produce children. For example, black partners tend to be positively selected with regard to education into Black/White marriages while white partners are negatively selected (CITATION). According to status exchange theory (CITATIONS), this pattern may be common to all types of interracial marriages within an unequal status hierarchy, although the evidence of status exchange for other forms of marriages is weaker (CITATION). How this affects the relative resources of biracial individuals will depend then in some part on the relative strength of these phenomena for a given pairing, the gender balance in that pairing, and how the gender of parents affects outcomes.
